---
title: "Facet Theory and Partitions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Facet Theory and Partitions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,      # Width in inches
  fig.height = 4,     # Height in inches
  dpi = 150,          # Resolution (DPI)
  dev = "png"         # Use PNG for sharp rendering
)
```

## Facets, Configurations, and Partitions

Facet Theory (FT) is a meta-approach to empirical research proposed by Louis Guttman ( ) and since then developed by many others ( ). We strongly advise to acquire a basic understanding of FT before using this package, for example Shye (1998). An essential step in FT applications is Multidimensional Scaling (MDS) of similarities/distances; for introductions see Borg and Groenen (2005), and also refer to the vignettes of the *smacof* package, to which this package is closely related and which should be installed when using *facpart*.

Assume you are studying some domain (e. g., intelligence), and you measure a set of items from that domain, say intelligence test items. Theoretically, each item is classified according to the elements of a *facet*: With respect to intelligence items, a typical facet is the *modality* of the test item, with elements being *verbal*, *numerical*, and *figural*. Thus, each test item is mapped onto exactly one of these categories. The theory is assumed to be confirmed if these elements can be identified as geometrical patterns in an empirical space of the test items.

The canonical approach is to submit the correlations among the items to a multidimensional scaling procedure. As a result we obtain 2-dimensional (or higher dimensional) configurations of points (= items), with highly correlated items being close to each other, and items with small or negative correlations being farther apart. This configuration serves as the starting point to search for patterns according to the theoretical classification from the faceted definition. A simple pattern might be that all items belonging to a specific category (e. g., verbal items) are in close vicinity, but apart from items of a different category. If no such separation can be detected, the theory must be modified.

The patterns emphasized by FT are called *regional partitions*. Three types of partitions are commonly identified:

- *Axial* partition: categories are separated by straight parallel lines.

- *Radial* partition: categories are separated by nested circles.

- *Angular* partition: categories are separated by lines with different angles from a common origin.

This will become clear from the examples below. The *facpart* package provides functions to search for best fitting partitions.

## The Basic Workflow

As an example we take the Intelligence study (Guttman & Levy, 1991). The correlation matrix of 12 intelligence test items and the assignment to two facets (Material, Modality) is provided in the *gutt91* data file (slightly modified from Guttman1991 from the smacof package). We extract the correlation matrix and the facet variables.

```{r gutt91}
library(facpart)

Kor <- gutt91$gutt91_cor
Facets <- gutt91$gutt91_var
round(Kor, 2)
Facets

```

In the Facets dataframe we can see that the first facet *Material* assigns to each item the materiality of the task the item represents: It can be 'oral' if it is just based on oral communication, or 'manual' if the participant has to actually do something with his or her hands, or 'paper' if the item requires to read/write from a sheet of paper. The second facet *Modality* defines the symbol system involved: 'verbal' refers to text in natural language, 'numerical' refers to numbers and mathematical notation, and 'figural' refers to drawings, geometric shapes, etc.

Theoretically, we expect that these distinctions show up in specific patterns or *partitions* of a geometrical representation of the items. Thus, we perform a multidimensional scaling analysis of the correlation matrix. Since MDS takes distances as input data, we have to convert the correlation matrix (which are similarities) to distances. The conversion of correlations to distances follows the equation $$d = \sqrt {1 - r}$$

The distances are then submitted to MDS. For converting and for MDS we use functions from the *smacof* package.

```{r MDS}
Kor_D <- smacof::sim2diss(Kor, method = "corr", to.dist = TRUE)

gutt91_mds <- smacof::mds(Kor_D, type = "ordinal")
gutt91_mds
```

The Stress-1 value of the resulting 2-dimensional (default) configuration is *0.102* which is usually considered a good fit. To plot the fit we can simply take the output object *gutt91_mds* as argument for the plot function. However, usually another approach is preferable which gives more control; in particular, we want the facet elements as labels, not the item names. Thus, we first plot only the coordinates of the 12 items, which are in *gutt91_mds\$conf*. Then we add the facet labels for *Material* with the text() function; the labels are in the Facets data frame. The resulting plot now nicely shows the configuration of the three categories for materiality, and we can check if a regular pattern can be identified.

```{r MDSplot}
plot(gutt91_mds)

plot(gutt91_mds$conf, 
     asp = 1, las = 1)
text(gutt91_mds$conf, labels = Facets$Material, 
     cex = 0.8, pos = 3)
abline(v=0); abline(h=0)
```

Let's assume that our theory predicts that the Material facet can be partitioned according to an *angular partition*, that is, from a common origin, lines or rays of different angles divide the space into wedge-like regions. The `angularPartition()` does exactly this: It searches for a best fitting angular partition, so that each region contains only items of one specific facet element. If that is not perfectly possible, it minimizes the number of misclassified items and yields the best possible partitioning. The necessary arguments are the point coordinates (crd =), and the facet labels for each point (group =). The coordinates must be a numeric matrix or data frame with two columns, and the facet labels must be a factor with levels corresponding to the labels.

```{r angularPartition1}
plot(gutt91_mds$conf, 
     asp = 1, las = 1)
text(gutt91_mds$conf, labels = Facets$Material, 
     cex = 0.8, pos = 3)

angularPartition(crd = gutt91_mds$conf,
                 group = Facets$Material, add = TRUE)
```

To summarize, the basic workflow is:

- Construct a correlation matrix K or distance matrix D of a set of n items. If necessare, convert K to D with \`smacof::sim2diss()\`.

- Construct facet labels: Assign to each of n items exactly one of k (k \< n) labels corresponding to their facet elements. Store the assignment as a factor F with length = n (same item order as in D).

- Apply multidimensional scaling to D, such as \`smacof::mds(D, ndim = 2, type = "ordinal")\`. Check for fit and store the result object as *out_mds*.

- Plot the point configuration, for example \`plot(out_mds)\`.

- Apply one of the partition functions; for example, \`angularPartition(crd = out_mds\$conf, group = F)\`. Necessary arguments are crd (the point coordinates) and group (the facet assignments).

- Check the resulting plot.

More details will be provided in the following discussion of the partition functions.
